Pillar 1 - Sequential Pattern Mining

### 2019-01-28 Pillar 1 - Sequential Pattern Mining.Rmd
delete conflict

Function that returns title of a course given its code.
```{r}
find_course <- function(code){

@@

```{r AR data prep}
d_transactions <- d_course %>%

  # Exclude 
  filter(
    Type != "Mandatory",              # (i) mandatory courses e.g. COR, CAP, etc
    ! Letters %in% c("SKI", "PRO",    # (ii) skills & projects (taken by majority of students)
                     "SAS", "SAH", "SAC") # (iii) courses of semester abroad
    ) %>%

  # Join with transcripts.
  left_join(d_transcript, by = "Course ID") %>%

  # Create transactions (student level)
  group_by(`Student ID`) %>%
    summarize(list_course = list(`Course ID`)) %>%
  ungroup

transactions <- as(d_transactions$list_course, "transactions")
```

## Apriori Algorithm
We apply Apriori algorithm:
```{r}
results <- apriori(
  data = transactions, 
  parameter = list(
    supp = 0.02,
    smax = 1, # max support
    conf = 0.4,
    minlen = 2,
    maxlen = 2
    ),
  appearance = NULL,
  control = NULL
  ) %>%
  inspect

colnames(results)[2] <- "association"

results %>%
  arrange(desc(confidence))
```
Now we include data for pass and fail:
```{r}
d_transaction_passed <- d_course %>%
  select(Type,`Course ID`, Letters) %>%
  filter(Type == "Elective",
         ! Letters %in% c("SKI", "PRO"),
         ! Letters %in% c("SAS", "SAH", "SAC")) %>%
  select(`Course ID`) %>%
  left_join(d_transcript, by = "Course ID") %>%
  mutate(Outcome = case_when(Grade=="NG"~"Fail",
                             Grade > 5.5 ~"Pass",
                             T~"Fail")) %>%
  mutate(Item= paste(`Course ID`, Outcome, sep = "_")) %>%
  select(`Student ID`, Item) %>%
  group_by(`Student ID`) %>%
  summarize(list_course = list(Item)) %>%
  group_by()

transactions_passed <- as(d_transaction_passed$list_course, "transactions")

```
We apply Apriori algorithm:
```{r}
results_passed <- apriori(
  data = transactions_passed, 
  parameter = list(
    supp = 0,
    smax = 1, # max support
    conf = 0.4,
    minlen = 2,
    maxlen = 2
    ),
  appearance = NULL,
  control = NULL
  ) %>%
  inspect

colnames(results_passed)[2] <- "association"

results_passed %>%
  separate(lhs, into = c("LHS", "LHS_pass"), sep= "_", remove = F) %>%
  separate(rhs, into = c("RHS", "RHS_pass"), sep= "_", remove = F) %>%
  filter(RHS_pass == "Fail}") %>%
  arrange(desc(count))
```

>>>>

Function that returns title of a course given its code.
```{r function find_course}
find_course <- function(code){

@@

```{r AR data prep 1}
d_transactions <- d_course %>%

  # Exclude 
  filter(
    Type != "Mandatory",                  # (i) mandatory courses e.g. COR, CAP, etc
    ! Letters %in% c("SKI", "PRO",        # (ii) skills & projects (taken by majority of students)
                     "SAS", "SAH", "SAC") # (iii) courses of semester abroad (uninformative)
    ) %>%

  # Join with transcripts.
  inner_join(d_transcript, by = "Course ID")
```

```{r AR data prep 2}
# Create transactions (student level)
transactions <- d_transactions %>%
  group_by(`Student ID`) %>%
    summarize(list_course = list(`Course ID`)) %>%
  ungroup

transactions <- as(transactions$list_course, "transactions")
```

```{r AR data prep 3 (PF)}
d_transactions_PF <- d_transactions %>%

  # Take outcome of course into account (pass or fail)
  mutate(PF = case_when(Grade <  5.5 ~ "fail",
                        Grade >= 5.5 ~ "pass"),
         `Course ID` = paste(`Course ID`, PF, sep = "_"))

  # Create transactions with pass/fail
transactions_PF <- d_transactions_PF %>%
  group_by(`Student ID`) %>%
    summarize(list_course = list(`Course ID`)) %>%
  ungroup

transactions_PF <- as(transactions_PF$list_course, "transactions")
```

## Apriori Algorithm
```{r}
clean_AR <- function(AR){

  AR %>%
    as("data.frame") %>%
    filter(
      count >= 1
    ) %>%
    mutate(
      rules = str_remove(rules, pattern = "_fail"),
      lhs = substr(rules, 2, 8),
      rhs = substr(rules, 15, 21),
      `rhs.support` = confidence / lift
      ) %>%
    mutate_at(
      c("support", "lhs.support", "confidence", "rhs.support", "lift"),
      funs(round(., 5))
      ) %>%
    select(
      lhs, rhs,
      support, count, lhs.support,
      confidence, rhs.support,
      lift
      ) %>%
    arrange(
      desc(count)
    )

}
```

We apply Apriori algorithm:
```{r AR apriori}
AR <- apriori(
  data = transactions, 
  parameter = list(
    supp = 0,    # min support
    smax = 1,    # max support
    conf = 0,    # min confidence
    minlen = 2,  # min length of rule
    maxlen = 2,  # max length of rule
    ext = TRUE,
    arem = "diff",
    aval = TRUE,
    minval = 0
    ),
  appearance = NULL,
  control = list(
    verbose = FALSE
    )
  ) %>%
  clean_AR

Now we include data for pass and fail:

```{r apriori PF}

# creating vector of fail course
course_id_fail <- d_transactions_PF %>%
  filter(PF == "fail") %>%
  distinct(`Course ID`)

# Apply apriori
AR_PF <- apriori(
  data = transactions_PF, 
  parameter = list(
    supp = 0, # min support
    smax = 1,    # max support
    conf = 0,    # min confidence
    minlen = 2,  # min length of rule
    maxlen = 2,  # max length of rule
    ext = TRUE,
    arem = "diff",
    aval = TRUE,
    minval = 0
    ),
  appearance = list(
    both = course_id_fail$`Course ID` # only include rules with fail on LHS and RHS
    ),
  control = list(
    verbose = FALSE
    )
  ) %>%
  clean_AR
```

```{r AR exploration}

# TODO: MAKE APP

data_rules <- AR
data_rules <- AR_PF

# Best support
data_rules %>%
  arrange(desc(support))

# Best confidence
data_rules %>%
  filter(count >= 5) %>%
  arrange(desc(confidence))

# Best lift
data_rules %>%
  filter(count >= 5) %>%
  arrange(desc(lift))
```
